Controlled Experiments for Word Embeddings

### 2015-09-04 article/main.tex
moderate adriaan's modifications

\begin{section}{Introduction}
The field of natural language processing (NLP) is presently undergoing rapid
developments, not in the least triggered by the tremendous commercial
promises it holds.  A timely research topic in NLP are word embeddings,
where words from a vocabulary are represented by dense, real-valued
vectors.  Instead of one-hot vectors that merely indicate the location
of a word in the vocabulary, dense
vectors of dimension much smaller than the vocabulary size are
constructed such that they carry syntactic and semantic information.
Irrespective of the technique chosen, word embeddings are
typically derived from word co-occurrences.  More specifically, in a
machine-learning setting, word embeddings are typically trained by
scanning a short window over all the text in a corpus.  This process can
be seen as sampling word co-occurrence distributions, where it is
recalled that the co-occurrence distribution of a target word $w$
denotes the conditional probability $\p(w'|w)$ that a word $w'$ occurs in
its context, i.e. given that $w$ occurred.  Most applications of word
embeddings explore not the word vectors themselves, but relations
between them to solve, for example, similarity and word relation tasks
\cite{vecchi-baroni-zamparelli2011}.  For these tasks, it was found that using normalised word
vectors improves performance.  Word vector length is therefore typically
ignored.

@@

\begin{section}{Related work}\label{related-work}
Our notion of word significance is reminiscent of the concept of
``contextual distinctiveness'' introduced earlier by McDonald and
Shillcock \cite{mcdonald2001contextual}, who also based their work on word co-occurrences.
They define contextual distinctiveness of a target word $w$ as the
Kullback-Leibler divergence $\mathrm{D} \big(\p(w') ||\p(w'|w) \big)$
between the distribution of context words, $\p(w')$, and the
co-occurrence distributions of that target word $w$.  Specifically,
\begin{equation} 
\label{cd}
\mathrm{D} \big(\p(w') ||\p(w'|w) \big) = \sum_{w'} \p(w'|w) \log_2
\frac{\p(w'|w)}{\p(w')},
\end{equation} 
where the sum is over the set of \textcolor{blue}{content?} words $w'$.
As pointed out by the
authors, contextual distinctiveness is a measure of the amount of
information provided by word $w$ about its contexts of use.  More specifically, if a
word $w$ tends to appear in a wide variety of contexts, the distribution
of the words it occurs with, i.e. the posterior distribution $\p(w'|w)$,
tends to diverge less from the prior distribution $\p(w')$, and its
contextual distinctiveness is accordingly small.  If, on the other hand,
$w$ typically appears in only a small number of different contexts, the
posterior distribution tends to diverge more from the prior, and a
larger value of contextual distinctiveness results.

@@

\begin{subsection}{Tokens derived from an artificial, meaningless word}\label{WFVEmeaningless}
Whereas the tokens introduced above all replace an existing word that
carries a meaning, we now include for comparison a high-frequency,
meaningless word.  We choose to introduce an artificial, entirely
meaningless word \word{VOID} into the corpus, rather than choose an
existing (stop)word whose meaninglessness is only supposed.  To achieve
this, we intersperse the word uniformly at random throughout the corpus so
that its global frequency is $0.005$.  The co-occurrence
distribution of \word{VOID} thus coincides with the global word
distribution.  The replacement procedure is then performed for this
word, using the same values for $p$ and $n$ as above.
Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of
these modifications on a sample text, where a higher global frequency of $0.05$ is used instead
for illustrative purposes.  Although the probability of inserting token
\word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
times that of inserting token \word{VOID\_3}, these three tokens all
appear once in this small piece of text.

\begin{figure}
	\begin{mdframed}
	\input{word-frequency-experiment-text-void.tex}
	\end{mdframed}
	\caption{The same example sentences as in
          Figure \ref{fig:word-frequency-experiment-text-cat} where
          instead of the word \word{cat} now the meaningless word
          \word{VOID} is replaced with tokens \word{VOID\_i}.  For
          illustrative purposes, the meaningless word \word{VOID} was
          here interspersed with a global frequency of $0.05$.}
	\label{fig:word-frequency-experiment-text-void}
\end{figure}
\end{subsection}

@@

%When read from right to left, this figure can be reinterpreted and
%related to the corresponding figure in
%Figure \ref{fig:word-frequency-experiment-graph} as follows.
If the average effect of co-occurrence noise were neglible, then the horizontal axis could be reinterpreted as
denoting term frequency and (reading the graph from right to left).
Thus a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
word vector length increases more or less linearly with term frequency.
This can not, however, be the complete picture, for the vectors
represented by the second synaptic layer, syn1neg, show a different
behavior in the two figures.  In particular, the length of these vectors
is seen from the right panel of Figure \ref{fig:co-occurrence-noise-graph}
to be completely insensitive to noise, i.e. frequency in our
reinterpretation, whereas the length of those shown in the right panel
of Figure \ref{fig:word-frequency-experiment-graph} does depend on
frequency.

>>>>

\begin{section}{Introduction}
Word embeddings represent words from the vocabulary as dense, real-valued
vectors.  Instead of one-hot vectors that merely indicate the location
of a word in the vocabulary, dense
vectors of dimension much smaller than the vocabulary size are
constructed such that they carry syntactic and semantic information.
Irrespective of the technique chosen, word embeddings are
typically derived from word co-occurrences.  More specifically, in a
machine-learning setting, word embeddings are typically trained by
scanning a short window over all the text in a corpus.  This process can
be seen as sampling word co-occurrence distributions, where it is
recalled that the co-occurrence distribution of a target word $w$
denotes the conditional probability $\p(w'|w)$ that a word $w'$ occurs in
its context, i.e. given that $w$ occurred.  Most applications of word
embeddings explore not the word vectors themselves, but relations
between them to solve, for example, similarity and word relation tasks
\cite{vecchi-baroni-zamparelli2011}.  For these tasks, it was found that using normalised word
vectors improves performance.  Word vector length is therefore typically
ignored.

@@

\begin{section}{Related work}\label{related-work}
Our notion of word significance is reminiscent of the concept of
``contextual distinctiveness'' introduced earlier by McDonald and
Shillcock \cite{mcdonald2001contextual}, who also based their work on word co-occurrences.
They define the contextual distinctiveness of a target word $w$ as the
Kullback-Leibler divergence
between the global word distribution $\p(w')$, and the
co-occurrence distribution $\p(w' | w)$ of the target word $w$.
So defined, contextual distinctiveness is a measure of the amount of
information provided by word $w$ about its contexts of use.  More specifically, if a
word $w$ tends to appear in a wide variety of contexts, the distribution
of the words it occurs with, i.e. the posterior distribution $\p(w' | w)$,
tends to diverge less from the prior distribution $\p(w')$, and its
contextual distinctiveness is accordingly small.  If, on the other hand,
$w$ typically appears in only a small number of different contexts, the
posterior distribution tends to diverge more from the prior, and a
larger value of contextual distinctiveness results.

@@

\begin{subsection}{Tokens derived from an artificial, meaningless word}\label{WFVEmeaningless}
Whereas the tokens introduced above all replace an existing word that
carries a meaning, we now include for comparison a high-frequency,
meaningless word.  We choose to introduce an artificial, entirely
meaningless word \word{VOID} into the corpus, rather than choose an
existing (stop)word whose meaninglessness is only supposed.  To achieve
this, we intersperse the word uniformly at random throughout the corpus so
that its global occurrence rate is $0.005$.  The co-occurrence
distribution of \word{VOID} thus coincides with the global word
distribution.  The replacement procedure is then performed for this
word, using the same values for $p$ and $n$ as above.
Figure \ref{fig:word-frequency-experiment-text-void} shows the effect of
these modifications on a sample text, where a higher word occurrence rate of $0.05$ is used instead
for illustrative purposes.  Although the probability of inserting token
\word{VOID\_1} is twice that of inserting token \word{VOID\_2} and four
times that of inserting token \word{VOID\_3}, these three tokens all
appear once in this small piece of text.

\begin{figure}
	\begin{mdframed}
	\input{word-frequency-experiment-text-void.tex}
	\end{mdframed}
	\caption{The same example sentences as in
          Figure \ref{fig:word-frequency-experiment-text-cat} where
          instead of the word \word{cat} now the meaningless word
          \word{VOID} is replaced with tokens \word{VOID\_i}.  For
          illustrative purposes, the meaningless word \word{VOID} was
          here interspersed with a occurrence rate of $0.05$.}
	\label{fig:word-frequency-experiment-text-void}
\end{figure}
\end{subsection}

@@

%When read from right to left, this figure can be reinterpreted and
%related to the corresponding figure in
%Figure \ref{fig:word-frequency-experiment-graph} as follows.
If the average effect of co-occurrence noise were neglible, then the horizontal axis of Figure \ref{fig:co-occurrence-noise-graph} could be reinterpreted as
denoting word frequency by reading the graph from right to left.
Thus a similar picture to Figure \ref{fig:word-frequency-experiment-graph} would emerge, namely that the
word vector length increases more or less linearly with word frequency.
This can not, however, be the complete picture, for the vectors
represented by the second synaptic layer, syn1neg, show a different
behavior in the two figures.  In particular, the length of these vectors
is seen from the right panel of Figure \ref{fig:co-occurrence-noise-graph}
to be completely insensitive to noise, i.e. frequency in our
reinterpretation, whereas the length of those shown in the right panel
of Figure \ref{fig:word-frequency-experiment-graph} does depend on
frequency.
